---
title: "Run CLK1 status survival analyses"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
author: Ryan Corbett, Jo Lynne Rokita
date: 2023, 2024
params:
  plot_ci: TRUE
---

**Purpose:** 

Runs survival analysis models in histology groups with CLK1 status as a predictor

## Usage 

Uses a wrapper function (`survival_analysis`) from utils folder. 

## Setup

#### Packages and functions

Load packages, set directory paths and call setup script

```{r}
library(tidyverse)
library(survival)
library(ggpubr)
library(ggplot2)
library(data.table)

root_dir <- rprojroot::find_root(rprojroot::has_dir(".git"))

data_dir <- file.path(root_dir, "data")
analysis_dir <- file.path(root_dir, "analyses", "survival")
input_dir <- file.path(analysis_dir, "results")
results_dir <- file.path(analysis_dir, "results")

# If the input and results directories do not exist, create it
if (!dir.exists(results_dir)) {
  dir.create(results_dir, recursive = TRUE)
}

source(file.path(analysis_dir, "util", "survival_models.R"))
```

Set metadata path

```{r set paths}
histologies_file <- file.path(data_dir, "histologies.tsv")

metadata_file <- file.path(input_dir, "splicing_indices_with_survival.tsv")

clk1_psi_file <- file.path(root_dir, "analyses",
                           "CLK1-splicing_correlations", "results",
                           "clk1-splice-events-rmats.tsv")
```

Wrangle data
Append clk1 group to `metadata`

```{r}
histologies <- read_tsv(histologies_file, guess_max = 10000) 

meta <- read_tsv(metadata_file) %>%
  left_join(histologies %>% dplyr::select(Kids_First_Biospecimen_ID,
                                          match_id)) %>%
  dplyr::mutate(extent_of_tumor_resection = fct_relevel(extent_of_tumor_resection,
                                                        c("Gross/Near total resection",
                                                          "Partial resection",
                                                          "Biopsy only", "Not Reported",
                                                          "Unavailable")))

clk1_psi <- data.table::fread(clk1_psi_file) %>%
   dplyr::filter(geneSymbol=="CLK1",
                 exonStart_0base=="200860124", exonEnd=="200860215") %>%
  dplyr::rename('Kids_First_Biospecimen_ID'=sample_id,
                clk1_PSI = IncLevel1) %>%
  dplyr::select(Kids_First_Biospecimen_ID, clk1_PSI)

metadata_joined <- meta %>%
  left_join(clk1_psi) %>%
  filter(!is.na(clk1_PSI))

# Calculate mean and standard deviation of clk1_PSI
mean_clk1_PSI <- mean(metadata_joined$clk1_PSI, na.rm = TRUE)
one_sd <- sd(metadata_joined$clk1_PSI, na.rm = TRUE)
two_sd <- 2*sd(metadata_joined$clk1_PSI, na.rm = TRUE)

# Add a new column for High/Low classification
metadata <- metadata_joined %>%
  mutate(clk1_group = case_when(clk1_PSI > (mean_clk1_PSI + one_sd) ~ "High CLK1 Exon 4",
                                clk1_PSI < (mean_clk1_PSI - two_sd) ~ "Low CLK1 Exon 4",
                                TRUE ~ NA_character_)
)

# remove samples with low Ns
# Create a table of counts by Histology and clk1_group
histology_table <- as.data.frame(table(metadata$Histology, metadata$clk1_group))
colnames(histology_table) <- c("Histology", "clk1_group", "Count")

# Filter histologies with at least 5 samples in both "High" and "Low"
valid_histologies <- histology_table %>%
  filter(!is.na(clk1_group)) %>%
  group_by(Histology) %>%
  summarise(min_count = min(Count)) %>%
  filter(min_count >= 5) %>%
  pull(Histology)

metadata_subset <- metadata %>%
  dplyr::filter(Histology %in% valid_histologies,
                !is.na(clk1_group)) %>%
  dplyr::mutate(clk1_group = factor(clk1_group,
                                        c("Low CLK1 Exon 4", "High CLK1 Exon 4")))

```

Define histology groups to run survival analyses on 

```{r}
subtype_df <- read_tsv(file.path(analysis_dir, "results", "subtypes-for-survival.tsv")) %>%
  filter(hist %in% valid_histologies,
         subtype %in% metadata_subset$mol_sub_group) %>%
  write_tsv(file.path(analysis_dir, "results", "subtypes-for-survival-clk1.tsv"))

groups <- unique(metadata_subset$Histology)

# folder names
folder <- c("LGG", "HGG", "GNG", "MB")
names(folder) <- groups
  
# create names vector for output files
plot_names <- c("lgg", "hgg", "gng", "mb")
names(plot_names) <- groups
```

Loop through histology groups and perform the following: 

- Define samples with high and low CLK1 PSI
- Run Kaplan-Meier survival analysis and save output for plotting
- Run cox proportional hazard models, including SI burden as covariate along with subtype (when applicable) and extent of tumor resection (LGG and GNG only). Save output for plotting

```{r}
for (group in groups){
  
  results_dir <- file.path(analysis_dir, "results", folder[[group]])
  
  if (!dir.exists(results_dir)) {
    dir.create(results_dir)
      
    }
  
  # Define output files for KM and coxph models
  km_os_model_file <- file.path(results_dir, 
                                    glue::glue("logrank_{plot_names[group]}_OS_clk1_group.RDS"))
  
  coxph_os_model_file <- ifelse(group == "Low-grade glioma",
                             file.path(results_dir,
                                       glue::glue("cox_{plot_names[group]}_OS_additive_terms_subtype_resection_clk1_group.RDS")),
                             file.path(results_dir,
                                       glue::glue("cox_{plot_names[group]}_OS_additive_terms_subtype_resection_clk1_group.RDS"))
  )
  
  km_efs_model_file <- file.path(results_dir, 
                                    glue::glue("logrank_{plot_names[group]}_EFS_clk1_group.RDS"))
  
  coxph_efs_model_file <- ifelse(group == "Low-grade glioma",
                             file.path(results_dir,
                                       glue::glue("cox_{plot_names[group]}_EFS_additive_terms_subtype_resection_clk1_group.RDS")),
                             file.path(results_dir,
                                       glue::glue("cox_{plot_names[group]}_EFS_additive_terms_subtype_resection_clk1_group.RDS"))
  )
  
  output_file <- file.path(results_dir,
                                     glue::glue("{plot_names[group]}_with_clk1_status.tsv"))
  
  # Subset histology
  group_hist <- metadata_subset %>%
    filter(Histology == group)
  
  if (group == "Low-grade glioma"){
    group_hist <- group_hist %>%
     dplyr::mutate(mol_sub_group = factor(mol_sub_group)) %>% 
     dplyr::mutate(mol_sub_group = fct_relevel(mol_sub_group, c("Wildtype", "BRAF V600E", "BRAF fusion",
                                                                "Other alteration", "SEGA"
                                                                )))
  }
  
  if (group == "Other high-grade glioma"){
    group_hist <- group_hist %>%
     dplyr::mutate(mol_sub_group = factor(mol_sub_group)) %>% 
     dplyr::mutate(mol_sub_group = fct_relevel(mol_sub_group, c("HGG, H3 wildtype",
                                                                "HGG, H3 wildtype, TP53",
                                                                "DHG, H3 G35")))
  }
  
  #Define high and Low CLK1 Exon 4 samples in ONLY samples with survival data and resection (LGG/GNT only)
  
  if (group == "Low-grade glioma"){
    
    group_hist <- group_hist %>%
      filter(!is.na(mol_sub_group) & (!is.na(OS_days) | !is.na(EFS_days)) & !is.na(extent_of_tumor_resection) & !extent_of_tumor_resection %in% c("Unavailable", "Not Applicable"))
    
  } else {
    group_hist <- group_hist %>%
      filter(!is.na(OS_days) | !is.na(EFS_days) & !is.na(extent_of_tumor_resection) & !extent_of_tumor_resection %in% c("Unavailable", "Not Applicable"))
    
  }
  
  # Generate kaplan meier survival models for OS and EFS, and save outputs
  kap_os <- survival_analysis(
    metadata  = group_hist,
    ind_var = "clk1_group",
    test = "kap.meier",
    metadata_sample_col = "Kids_First_Biospecimen_ID",
    days_col = "OS_days",
    status_col = "OS_status"
  )

  readr::write_rds(kap_os, km_os_model_file)
  
  kap_efs <- survival_analysis(
    metadata  = group_hist,
    ind_var = "clk1_group",
    test = "kap.meier",
    metadata_sample_col = "Kids_First_Biospecimen_ID",
    days_col = "EFS_days",
    status_col = "EFS_status"
  )

  readr::write_rds(kap_efs, km_efs_model_file)

  # Run cox proprtional hazards models for OS, including appropriate covariates per histology
  add_model_os <- suppressWarnings(fit_save_model(group_hist,
                                  terms = ifelse(
                                      group == "Low-grade glioma",
                                      "mol_sub_group+extent_of_tumor_resection+clk1_group+age_at_diagnosis_days+age_at_diagnosis_days",
                                      ifelse(group == "Mixed neuronal-glial tumor",
                                             "mol_sub_group+extent_of_tumor_resection+clk1_group+age_at_diagnosis_days+age_at_diagnosis_days",
                                              "mol_sub_group+extent_of_tumor_resection+clk1_group+age_at_diagnosis_days"
                                  )),
             coxph_os_model_file,
             "multivariate",
             years_col = "OS_years",
             status_col = "OS_status"
            ))
  
  # Run cox proprtional hazards models for EFS, including appropriate covariates per histology
  add_model_efs <- suppressWarnings(fit_save_model(group_hist,
                                  terms = ifelse(
                                      group == "Low-grade glioma",
                                      "mol_sub_group+extent_of_tumor_resection+clk1_group+age_at_diagnosis_days+age_at_diagnosis_days",
                                      ifelse(group == "Mixed neuronal-glial tumor",
                                             "mol_sub_group+extent_of_tumor_resection+clk1_group+age_at_diagnosis_days+age_at_diagnosis_days",
                                              "mol_sub_group+extent_of_tumor_resection+clk1_group+age_at_diagnosis_days"
                                  )),
             coxph_efs_model_file,
             "multivariate",
             years_col = "EFS_years",
             status_col = "EFS_status"
            ))
  
  group_hist %>%
    filter(!is.na(clk1_group)) %>%
    select(-OS_years, -EFS_years, 
           -Kids_First_Participant_ID) %>%
    write_tsv(output_file)
  
}
```

Run same survival analyses on molecular subtypes with sufficient sample size

```{r}
# Loop through subtypes and run survival analyses
for (i in 1:nrow(subtype_df)) {
  
  results_dir <- file.path(analysis_dir, "results", subtype_df$foldername[i])
  
  # define output files
  km_os_model_file <- file.path(results_dir, 
                                    glue::glue("logrank_{subtype_df$names[i]}_{subtype_df$subtype_name[i]}_OS_clk1_group.RDS"))
  
  coxph_os_model_file <- ifelse(subtype_df$hist[i] == "Low-grade glioma",
                             file.path(results_dir,
                                      glue::glue("cox_{subtype_df$names[i]}_{subtype_df$subtype_name[i]}_OS_additive_terms_resection_clk1_group.RDS")),
                             file.path(results_dir,
                                       glue::glue("cox_{subtype_df$names[i]}_{subtype_df$subtype_name[i]}_OS_additive_terms_resection_clk1_group.RDS"))
  )
  
  km_efs_model_file <- file.path(results_dir, 
                                    glue::glue("logrank_{subtype_df$names[i]}_{subtype_df$subtype_name[i]}_EFS_clk1_group.RDS"))
  
  coxph_efs_model_file <- ifelse(subtype_df$hist[i] == "Low-grade glioma",
                             file.path(results_dir,
                                       glue::glue("cox_{subtype_df$names[i]}_{subtype_df$subtype_name[i]}_EFS_additive_terms_resection_clk1_group.RDS")),
                             file.path(results_dir,
                                       glue::glue("cox_{subtype_df$names[i]}_{subtype_df$subtype_name[i]}_EFS_additive_terms_resection_clk1_group.RDS"))
  )
  
  SI_burden_output_file <- file.path(results_dir,
                                     glue::glue("{subtype_df$names[i]}_{subtype_df$subtype_name[i]}_with_clk1_group.tsv"))
  
    #Subset metadata for histology subtype
    subtype_hist <- metadata_subset %>%
      filter(Histology == subtype_df$hist[i]) #& grepl(subtype_df$subtype[i], mol_sub_group))
    
    if(subtype_df$hist[i] %in% c("Mixed neuronal-glial tumor", "Low-grade glioma"))
      {
      subtype_hist <- subtype_hist %>%
        dplyr::filter(!is.na(extent_of_tumor_resection) & !extent_of_tumor_resection %in% c("Not Applicable", "Unavailable") & (!is.na(OS_days) | !is.na(EFS_days)))
      
    } else {
      
      subtype_hist <- subtype_hist %>%
        dplyr::filter((!is.na(OS_days) | !is.na(EFS_days)))
    }
    
    #Run kaplan meier survival models and save to output
  kap_os <- survival_analysis(
              metadata  = subtype_hist,
              ind_var = "clk1_group",
              test = "kap.meier",
              metadata_sample_col = "Kids_First_Biospecimen_ID",
              days_col = "OS_days",
              status_col = "OS_status"
              )

  readr::write_rds(kap_os, km_os_model_file)
  
  kap_efs <- survival_analysis(
                metadata  = subtype_hist,
                ind_var = "clk1_group",
                test = "kap.meier",
                metadata_sample_col = "Kids_First_Biospecimen_ID",
                days_col = "EFS_days",
                status_col = "EFS_status"
                )
  
  readr::write_rds(kap_efs, km_efs_model_file)
    
    # Run coxph models and save to output
  add_model_os <- suppressWarnings(fit_save_model(subtype_hist,
                                  terms = "extent_of_tumor_resection+clk1_group+age_at_diagnosis_days",
                                  coxph_os_model_file,
                                  "multivariate",
                                  years_col = "OS_years",
                                  status_col = "OS_status")
                                 )
  
  add_model_efs <- suppressWarnings(fit_save_model(subtype_hist,
                                   terms = "extent_of_tumor_resection+clk1_group+age_at_diagnosis_days",
                                   coxph_efs_model_file,
                                   "multivariate",
                                   years_col = "EFS_years",
                                   status_col = "EFS_status"
                                  ))
  subtype_hist %>%
    filter(!is.na(clk1_group)) %>%
    select(-OS_years, -EFS_years,
           -Kids_First_Participant_ID) %>%
    write_tsv(SI_burden_output_file)
    
}
```